---
title: "R Camp 2019 HW3"
author: "Alex Pugh"
date: "8/6/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clear environment


```

## Problem 1

Unilaterally taking a dispute to court is not generally viewed as a friendly act between states. I am interested in the impact of a unilateral appeal to the ICJ on a population's percetion of another state. For a case to go to the court, there must already exist some sort of dispute. Presumably, the presence of a dispute sours the perception of the other state in the mind of the population. Does the unilateral appeal by State A against State B have an additional negative impact on the population of State B's perception of State A? Or, does a unilateral appeal to the court make the population of State B percieve State A as being more willing to settle the dispute and find a peaceful solution? 

To explore this question, I will simulate an experiment. I will make State B the United States, so the population of my experiment would be the adult population of the United States. Relevant characteristics of the population that I will simulate are age, gender, and political affiliation. I will also consider the state-level percentage of bachelor's degree holder. I expect that younger people, women, Democrats, and people in states with higher educational attainment will generally have more positive perceptions of the other state than others because they may be more open to the international community and have fewer nationalist tendencies. They may be more likely to recognize that the US having a dispute with another state does not make the other state a "bad" state or an enemy. 

In my experiment, I would inform people that it is 2025 and the US and Canada are in a dispute, potentially over the maritime boundary in the north east given changes in fish migration patterns. I would provide a few examples of the dispute in action and the tensions betweeen the countries. Those in the control would then be asked to rank on a scale ranging from 0 to 10 how favorable they view Canada. Those in the treatment group would be told that Canada had filed a unilateral application to the International Court of Justice and that the issue would go before the court. Then they would be asked a few question and will get a perception score on a scale ranging from 0 to 10 how favorable they view Canada. I expect that the treatment will have a negative effect on their perception and resulting score.  
First I will create the population data for the covariates. For computational ease, I have set the adult population of the US to 100 million, with approximately 50% of women. Once we have the individual attributes, we can create a perception score (DV) for if the individual is in the tratment group and if the individual is not in the treatment group. Once we sample and assign individuals to the treatment group, we can then use the appropriate perception score as the observed score.


```{r}
#Adult population of the US, set as 100 million 
simN <- 100000000

#set seed
set.seed(111222)
#generate gender variable
female <- rbinom(n=simN, size=1, prob=0.50) #50 percent men and women 

head(female)


#set seed
set.seed(111222)
#create age variable ranging from 18+
#use normal distribution and add
age <-  rpois(n=simN, lambda=38) 
summary(age)
#add 10 to all values under 18
age <- ifelse(age < 18, age+10, age)
summary(age)

#set seed
set.seed(111222)
#create political affiliation
p1 <- rep(NA, simN)
p2 <- ifelse(age <= 35 & runif(simN)<0.65, 1, p1)    #make 65% of those under 35 democrat
p3 <- ifelse(age >30 & female ==1 & runif(simN)<0.60, 1, p2) # make 60% of women over 30 Democrat
p4 <- ifelse(age >30 & female ==0 & runif(simN)<0.45, 1, p3) # make 45% of men over 30 Democrat
democrat <- ifelse(is.na(p4)==T, 0, 1)
head(democrat)

#set seed
set.seed(111222)
# create state level % bachelor's degree holders
edu <- rnorm(n=50, mean=27, sd=1.5)
edu <- round(edu, digits = 3)
summary(edu)

#set seed
set.seed(111222)
state.edu <- sample(edu, size = simN, replace = TRUE)
#hist(state.edu)
#summary(state.edu)
#make state.edu a decimal
state.edu <- state.edu/100

# create subject number
respondent <- seq(from=1, simN, by=1)

# simulate treated and untreated outcome variable
# set coefficients
 a <- 6 
 b.age      <- -0.08    
 b.female    <- 1.1 
 b.democrat <- 0.7
 b.state.edu <- 1.0
 tau <- -1.2 #set treatment effect

# set seed and create error term
set.seed(111)
error <- rnorm(n=simN, mean=0, sd=0.17)
summary(error)

#create dataframe
sim.dat <- data.frame(respondent, age, female, democrat, state.edu) 

#create untreated and treated outcome
sim.dat$Y_0 <- a + tau*0 + b.age*age + b.female*female + b.democrat*democrat + b.state.edu+state.edu + error  # Potential outcome when not treated
sim.dat$Y_1 <- a + tau*1 + b.age*age + b.female*female + b.democrat*democrat + b.state.edu+state.edu + error  # Potential outcome when treated

#check range and distribution of treated and untreated outcomes
summary(sim.dat$Y_0)
summary(sim.dat$Y_1)

#check data
head(sim.dat, n=4)
unique(sim.dat$state.edu) #check to ensure 50 values

# DONE WITH POPULATION DATA

```

Next, we will randomly sample 50 individuals from the population. Then we will randomly assign then to the treatment and control group and explore the difference in means between the two groups. 

```{r}

#set seed
set.seed(111222)
#sample size
N_samp <- 50
#get sample of 50 names
index <- sample(1:nrow(sim.dat), size = N_samp, replace =FALSE)
index
samp.exp <- sim.dat[index,]


#create binary treatment variable
#set seed
set.seed(111)
samp.exp$d <- rbinom(n=nrow(samp.exp), size=1, prob=0.5) 
#look at distribution
table(samp.exp$d)

#create observed outcome
samp.exp$Y_obs <- samp.exp$d*samp.exp$Y_1 + (1-samp.exp$d)*samp.exp$Y_0  # Observed outcomes
# Y_obs = d*Y_1 + (1-d)*Y_0
samp.exp$Status <- ifelse(samp.exp$d==1, "Treated", "Control")      # Copying the treatment status into sample data


#Visualize with boxplot
boxplot(samp.exp$Y_obs[samp.exp$Status=="Treated"],
        samp.exp$Y_obs[samp.exp$Status=="Control"],
        names=c("Treated", "Control"), ylab="Perception Score",
        main="Distribution of Perception Scores\nN = 50",
        col=c("light blue", "blue"))


#Create plot of mean perception for each group
t <- c(0,1)
y_mean <- c(mean(samp.exp$Y_obs[samp.exp$Status=="Control"]),
            mean(samp.exp$Y_obs[samp.exp$Status=="Treated"]))
y_sdv <- c(sd(samp.exp$Y_obs[samp.exp$Status=="Control"]),
           sd(samp.exp$Y_obs[samp.exp$Status=="Treated"]))

plot(y_mean ~ t, pch=19, col="red", ylim=range(c(y_mean-y_sdv, y_mean+y_sdv)),
     xlab="Treatment Status", ylab="Y_obs ± sd",xaxt="n") # Without x-axis lable
axis(1, at = seq(00, 1, by = 1), las=1) # If las=2, numbers will be flipped by 90 degree
arrows(t, y_mean-y_sdv, t, y_mean+y_sdv, length=0, angle=90, lwd=3)
title("Simulated Result")


#perform a T-test (difference in means test)
t.test(Y_obs ~ Status, data=samp.exp)


# Estimating the average treatment effect
tau_hat <- mean(samp.exp$Y_obs[samp.exp$Status=="Treated"]) -
           mean(samp.exp$Y_obs[samp.exp$Status=="Control"])
tau_hat

# Population level treatment (fixed=constant) effect
tau_true <- mean(sim.dat$Y_1 - sim.dat$Y_0) # Mean Difference of Individual Potential Outcomes
tau_true
tau_true == tau


```

With the sample of 50 individuals, I was not able to perfectly recover the treatment effects. Instead, I over-estimated the treatment effect. Below are graphs of the distribution of the covariates across the treatment and control groups. The control group had more women and more democrats, which likely inflated the estimate of the treatment effect as the coefficients for female and democrat were positive.

```{r}

#Compare covariates between treatment and control
#age
par(mfrow=c(1,2))  
hist(samp.exp$age[samp.exp$Status=="Treated"], breaks = 6, main = "Distribution of Age\nAmong Treatment", xlab="Age")
hist(samp.exp$age[samp.exp$Status=="Control"], breaks = 6, main = "Distribution of Age\nAmong Control", xlab="Age")

#state.edu
par(mfrow=c(1,2))  
hist(samp.exp$state.edu[samp.exp$Status=="Treated"], breaks = 6, main = "Distribution of State Education\nAmong Treatment", xlab="State % W/ Bachelor's")
hist(samp.exp$state.edu[samp.exp$Status=="Control"], breaks = 6, main = "Distribution of State Education\nAmong Control", xlab="State % W/ Bachelor's")

#female
par(mfrow=c(1,2))  
barplot(table(samp.exp$female[samp.exp$Status=="Treated"]), main = "Distribution of Gender\nAmong Treatment", xlab="Gender", names.arg = c("Men", "Women"), col = c("light blue","pink"))
barplot(table(samp.exp$female[samp.exp$Status=="Control"]), main = "Distribution of Gender\nAmong Control", xlab="Gender", names.arg = c("Men", "Women"), col = c("light blue","pink"))

#democrat
par(mfrow=c(1,2))  
barplot(table(samp.exp$democrat[samp.exp$Status=="Treated"]), main = "Distribution of Political Affiliation\nAmong Treatment", xlab="Political Affiliation", names.arg = c("Republican", "Democrat"),
        col = c("red","blue"))
barplot(table(samp.exp$democrat[samp.exp$Status=="Control"]), main = "Distribution of Political Affiliation\nAmong Control", xlab="Political Affiliation", names.arg = c("Republican", "Democrat"),
        col = c("red","blue"))

```

## Problem 2

Next, we repeat the same simulation, changing the number of samples from 50, 100, 150,200 to 250, and create a graph that combines five plots for the difference in distributions. From these plots, we can see that there is consistently a negative treatment effect. 

```{r}
#create vector of sample sizes
sam <- c(50,100,150,200,250)
#create parameters for graphs
par(mfrow=c(3,2))

#create empty vector for treatment effects
t1 <- rep(NA, 5)
#create vector for boolean TF if tau est is == tau true
t2 <- rep(NA, 5)

#create object for i in t1 
loop <- 1

#Loop through sample sizes and combine boxplots in one graph
for(i in sam){
 #set seed
set.seed(111)
#sample size
N_samp <- i
#get sample of names
index <- sample(1:nrow(sim.dat), size = N_samp, replace =FALSE)
samp.exp <- sim.dat[index,]


#create binary treatment variable
#set seed
set.seed(111)
samp.exp$d <- rbinom(n=nrow(samp.exp), size=1, prob=0.5) 
#look at distribution
table(samp.exp$d)

#create observed outcome
samp.exp$Y_obs <- samp.exp$d*samp.exp$Y_1 + (1-samp.exp$d)*samp.exp$Y_0  # Observed outcomes
# Y_obs = d*Y_1 + (1-d)*Y_0
samp.exp$Status <- ifelse(samp.exp$d==1, "Treated", "Control")      # Copying the treatment status into sample data

#Visualize with boxplot
boxplot(samp.exp$Y_obs[samp.exp$Status=="Treated"],
        samp.exp$Y_obs[samp.exp$Status=="Control"],
        names=c("Treated", "Control"), ylab="Perception Score",
        main=paste0("Distribution of Perception Scores\nN =", sep=" ", i),
        col=c("light blue", "blue")) 

#perform a T-test (difference in means test)
t.test(Y_obs ~ Status, data=samp.exp)


# Estimating the average treatment effect
tau_hat <- mean(samp.exp$Y_obs[samp.exp$Status=="Treated"]) -
           mean(samp.exp$Y_obs[samp.exp$Status=="Control"])
t1[loop]<- tau_hat

# Population level treatment (fixed=constant) effect
tau_true <- mean(sim.dat$Y_1 - sim.dat$Y_0) # Mean Difference of Individual Potential Outcomes
tau_true
t2[loop] <- tau_true == tau

#add one to loop
loop <- loop + 1
}
#View Boolean
t2

#plot 
par(mfrow=c(1,1))
hist(t1, xlim = c(-1.15,-1.35), main = "Distribution of Estimated Treatment Effect", 
     xlab = "Treatement Effect")
abline(v = tau, col="red", lwd=3, lty = 2)
legend("topright", legend="Mean" , lty = 2, lwd = 3, col = "red")
 
```

Even with increasing the sample size, I was not able to accurately recover the treatment effect. My estimated treatment effects were consistently larger in magnitude than the -1.2 true population treatment effect. As will previously, this was likely due to differences in distribution of individuals between the treatment and control, probably with more women/democrats in the control than in the treatment.   
